---
title: "R Notebook"
output: html_notebook
---

Read in the data

```{r}
africa <- read.table(text = "X,higher,percent,section
10,1,22,1
30,1,35,1
10,1,30,1
30,0,15,1
10,0,9,1
10,0,3,1
30,1,37,1
30,0,20,1
10,1,28,1
30,0,20,1
10,1,25,1
30,0,20,1
10,1,30,1
10,1,30,1
30,1,33,1
30,0,15,1
10,1,15,1
30,0,25,1
30,0,25,1
10,1,68,1
30,0,20,1
30,0,13,1
10,1,65,1
10,1,15,1
10,0,8,1
10,1,30,1
30,0,20,1
10,1,50,1
10,0,5,1
10,1,27,1
10,1,23,1
30,1,33,1
10,1,19,2
30,0,23,2
30,1,32,2
10,0,22,2
10,0,7,2
30,0,1,2
10,1,18,2
30,0,10,2
30,0,28,2
30,1,67,2
10,1,27,2
30,0,25,2
30,0,12.5,2
10,1,30,2
30,0,20,2
10,0,2,2
10,0,3,2
10,1,20,2
10,1,20,2
10,1,18,2
30,0,20,2
30,0,10,2
10,1,20,2
10,1,50,2
30,0,12,2
10,1,15,2
30,1,35,2
30,1,60,2
10,1,27,3
30,0,20,3
30,1,44,3
10,0,15,3
10,1,10,3
10,0,20,3
30,0,15,3
30,0,12,3
10,1,25,3
10,1,30,3
10,1,40,3
30,0,15,3
30,1,82,3
30,1,33,3
30,0,25,3
10,1,20,3
30,1,65,3
10,1,25,3
30,0,15,3
30,0,15,3
30,0,5,3
30,0,17,3
30,0,18,3
10,0,50,3
30,0,10,3
10,1,15,3
30,0,5,3
10,1,25,3
30,0,15,3
30,0,25,3
10,1,15,3
30,0,20,3
10,1,50,3
10,1,15,3
30,0,15,3
30,0,15,3
30,1,40,3
10,1,10,3
30,0,15,3
10,1,23,3
30,0,23,3
30,1,25,3
30,0,15,3
10,0,5,3
10,1,17,3
30,0,25,3
30,0,23,3
10,1,20,3
10,1,15,3", header = TRUE, sep = ",")
```


```{r}
africa$X <- factor(africa$X)
temphigher <- factor(africa$higher, levels=c(1,0), labels=c("Higher", "Lower"))

table(africa$higher, temphigher)

africa$higher <- temphigher
```


Create table
You should put the explanatory variable first (and the response second)

```{r}
africatab <- table(africa$X, africa$higher)
```


Make proportions table margins

```{r}
addmargins(africatab)
```


Calculate proportion percentages for each row and column

```{r}
prop.table(africatab, margin=1) # calculates row proportions -- what we want

prop.table(africatab, margin=2) # calculates proporitions within columns
```


Perform a binomial test (calculates confidence intervals)

```{r}
apply(africatab, 1, binom.test)
```


Calculate the p-value

```{r}
1-pchisq(29.86, 1)
```


Perform a chi square test with one function

```{r}
# this does the full calculation -- use this one
chisq.test(africatab, correct=FALSE)

# chisq test WITH continuity correction -- for very small sample sizes
chisq.test(africatab, correct=TRUE)
```


If our sample size is not large, we need to do Fisher's Exact Test (FET)

```{r}
fisher.test(africatab)
```


Calculate the CI for difference in proportions

```{r}
prop.test(africatab, correct = FALSE)
prop.test(africatab)
```

